Linear response theory and optimal control for a molecular system under nonequilibrium conditions

### 2013-08-23 writedown/linear.response/paper.tex
improves according to Carsten

@@

\subsubsection*{Girsanov transformation}
We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook \cite{oksendal2003stochastic}. To this end, let $x_{t}=x_{t}(\omega)$ and $X_{t}=X_{t}(\omega)$ be the solutions of our generic stochastic differential equations
\begin{subequations}\label{sde2}
\begin{align}
dx_{t} & = b(x_{t},t)dt + a(x_{t})dB_{t}\,,\quad 0\le t\le T  \label{sde2-1}\\
dX_{t} & = (b(X_{t},t) + \eps v_{t})dt +a(X_{t})dB_{t}\,,\quad 0\le t\le T \label{sde2-2}
\end{align}
\end{subequations}
for $T<\infty$ and deterministic initial conditions
\[
x_{0}(\omega) = X_{0}(\omega) = x\quad\textrm{(almost surely)}\,.
\]
Suppose that there exists an auxiliary stochastic process \redc{ $\xi_{t}\in\R^{n}$} such that 
\begin{equation}\label{xi}
a(X_{t})\xi_{t} = v_{t}\,.
\end{equation}
The auxiliary variable $\xi$ will be called \emph{control variable}. We define  
\[
W_{t} = \eps\int_{0}^{t}\xi_{s}\,ds + B_{t}\,,\quad 0\le t\le T\,,
\]
which allows us to rewrite (\ref{sde2-2}) as
\begin{equation}\label{sde2-3}
dX_{t} = b(X_{t},t)dt +a(X_{t})dW_{t} 
\end{equation}
It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003stochastic}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{stroock2006}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$. Specifically, let $P$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $Q$ on the space of continuous trajectories by 
\[
dQ=M_{T}dP
\]
with 
\begin{equation}\label{likelihood}
M_{t} = \exp\left(-\eps\int_{0}^{t}\xi_{s}\cdot dB_{s} - \frac{\eps^{2}}{2}\int_{0}^{t}|\xi_{s}|^{2}ds\right)\,,\quad 0\le t\le T\,.
\end{equation}
Technical details aside, the Girsanov theorem implies that $W_{t}$  for any functional 
\[
\varphi(\omega)=\int_{0}^{T}f((X_{s}(\omega))\,ds + g(X_{T}(\omega))
\] 
that is integrable \wrt $\nu$, we have the identity\footnote{One of the omitted technical details is Novikov's condition \cite[pp.~162]{oksendal2003stochastic} that guarantees that $(M_{t})_{0\le t\le T}$ is a Martingale, which by $\bE_{Q}[1]=\bE_{P}[M_{T}]=1$ implies normalizability of the new probability measure $Q$.}
\[
\bE_{P}[\varphi] := \int_{\Omega} \varphi(\omega)\, dP(\omega) = \int_{\Omega} \varphi(\omega) M_{T}^{-1}(\omega)\,dQ(\omega) =: \bE_{Q}[M^{-1}_{T}\varphi]\,.
\]
where 
\[
M_{T}^{-1} = \exp\left(\eps\int_{0}^{T}\xi_{s}\cdot dW_{s} - \frac{\eps^{2}}{2}\int_{0}^{T}|\xi_{s}|^{2}ds\right)
\]
is the density of $P$ relative to $Q$. Note that the expectation on the right hand side corresponds to the unperturbed dynamics, because $W_{t}$ is a standard Brownian motion under $Q$, and the expectation is over all realizations of (\ref{sde2-3}) starting from $X_{0}=x$. On the other hand, $X_{t}$ under $P$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).

@@

We are interested in the dynamics under a small perturbation $\delta u$ of the control $u$, Specifically, we assume that $v$ in (\ref{sde2}) is of the form 
\begin{equation}\label{eqn:langevin-1c}
v_{t} = c(X_{t})\delta u(t)
\end{equation}
so that equation (\ref{xi}) that determines the change of measure in terms of the auxiliary control variable $\xi$ (and thus the linear response) reads 
\[
\sqrt{2\beta^{-1}}\recheck{\gamma}^{1/2} \xi_{t} = D(X_{t})\,\delta u(t)\,.
\]
\recheck{Hence the equation for $\xi_{t}$ is solvable if and only if $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\gamma^{1/2})$, which, since $\gamma$ has full rank, means no restriction on $D(\cdot)$.}
  % means that $D(\cdot)$ must be invertible almost everywhere.}

@@

The nonequilibrium linear response formula can be used for solving certain optimal control problems. To this end, let us remain in the setting of equations (\ref{eqn:langevin-1d})--(\ref{eqn:neq-response}) and assume that we are interested in choosing the nonequilibrium forcing $u$, such that the expected value $I=\bE[\psi]$ of some utility function 
\[
\psi(u) = \int_{0}^{T}\left\{f(x_{s}) - \frac{1}{2}|u(s)|^{2}\right\}\,dt + g(x_{T}) 
\]
is maximized where $x_{t}$ is the solution to the controlled Langevin equation (\ref{eqn:langevin-1d}) for $\eps=0$; the functions $f$ and $g$ are the running cost and the terminal cost, which are assumed to be continuous and bounded from above; without loss of generality, $f$ and $g$ are assumed to depend only on the positions $q:=x_{1}$ and not on the momenta $p:=x_{2}$. The quadratic term is a penalization that makes sure that the optimal control remains bounded \cite{schuette2012,hartmann2012}. Let us moreover assume that the controls are open-loop (i.e. without feedback) and can be represented by 
\[
u(t) = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
\]
with suitably chosen time dependent, bounded and Lipschitz continuous vector fields $\Phi_k\colon[0,T]\to\R^{n}$, $k=1,\ldots,K$. %In principle we could even go for closed-loop (i.e.~feedback) controls by making $\Phi_{k}$ state dependent. 
Specifically, we want to solve the optimal control problem of the following standard form: 
\begin{equation}\label{ocp1}
\begin{aligned}
 \max_{u\in\cU} \,& I(u)\quad \textrm{s.t.}\\
dq_t & =  \nabla_p H(q_t,p_t)dt\\
dp_t & =  -\nabla_q H(q_t,p_t)dt - \gamma \nabla_p H(q_t,p_t)dt + D(q_t)u(t) dt+\sigma dB_t\,,
\end{aligned}
\end{equation}
\redc{with $(q_0, p_0)$ being subject to a given initial distribution $\mu$, and}
\begin{equation}\label{ocp2}
I(u) = \bE\left[\int_{0}^{T}\left\{f(q_{s}) - \frac{1}{2}|u(s)|^{2}\right\}\,dt + g(q_{T}) \right]\,,
\end{equation}
with the expectation being over all realizations of the controlled Langevin dynamics in (\ref{ocp1}) with \recheck{the} prescribed initial distribution. Here, the space of admissible controls $\cU$ consists of all bounded controls of the form 
\begin{equation}\label{admissible}
u(t) = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
\end{equation}

@@

In principle, optimal control problems such as (\ref{ocp1})--(\ref{ocp2}) can be solved by dynamic programming, i.e.~by solving the corresponding Hamilton-Jacobi-Bellman PDE \cite{fleming2006}. Except for very simple, essentially one-dimensional systems, solving Hamilton-Jacobi-Bellman equations is not an easy task, so we pursue a different strategy here. The idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and that we can do a gradient search in the unknown coefficients $a_{k}$, using the iteration
\[
u^{(n+1)} = u^{(n)} + \tau_{n}\nabla I(u^{(n)})\,,
\]
with $\tau_{n}>0$ being an adjustable parameter that determines the length of each gradient step. The gradient of $I$ can be easily evaluated using the linear response formula. To see this recall the notion of functional (G\^ateaux) derivatives as directional derivatives along a function $v$ (from a suitable function space): 
\begin{equation*}
\frac{\delta I}{\delta u} = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps v) = \langle\nabla I(u),v\rangle\,.
\end{equation*} 
For existence of the functional derivative \wrt general It\^o stochastic processes, we refer to \cite{bell2006}.  
Now the idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and do a gradient search in the unknown coefficients $a_{k}$. This requires to compute the gradient \wrt the coefficients. Let the vector  
\[
\delta u = \sum_{k=1}^K \delta a_k \Phi_k(t)
\]   
denote the direction along which we want to differentiate where $\delta a_{1},\ldots,\delta a_{K}$ are the coefficients of the vector $\delta u$ in the basis of the $\Phi_{k}$, and note that 
\begin{equation*}
\begin{aligned}
\frac{\delta I}{\delta u} & = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps \delta u)\\
& = \left(\lim_{\eps\to 0}\frac{I(u+\eps\delta u) - I(u)}{\eps}\right)\cdot \delta u\\
& = \left(\lim_{\eps\to 0}\frac{\bE_{P^{\eps}}[\psi(u+\eps \delta u)] - \bE_{P^{0}}[\psi(u)]}{\eps}\right)\cdot\delta u\\
\end{aligned}
\end{equation*} 
provided that the limit exists. The above iteration therefore is equivalent to
\begin{equation}\label{gd}
a_{k}^{(n+1)} = a_{k}^{(n)} + \tau_{n}\left.\frac{\partial I}{\partial a_{k}}\right|_{a_{k}=a_{k}^{(n)}}\,,
\end{equation}
with
\begin{equation}\label{dida}
\frac{\partial I}{\partial a_{k}} = \bE_{P^0}\left[\psi\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} + \int_{0}^{T}u(s)\cdot\Phi_{k}(s)\,ds \right] \,,
\end{equation}
and $\bE_{P^0}[\cdot]$ being the expectation over all realizations of (\ref{eqn:langevin-1d}) with $\eps=0$ and $u=u^{(n)}$ being the current iterate and arbitrary initial conditions $x=(q,p)$, e.g., 
\[
\bE_{P^{0}}[\psi] = \int_{\R^{2n}}\bE_{P^{0}_{x}}[\psi]\,d\recheck{\mu}(x) \,,
\]
with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition,
\redc{and $\mu$ is the prescribed initial distribution.}
Note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$. Further note that, in general, the updated coefficient $a^{(n+1)}$ and hence the updated control $u^{(n+1)}$ will depend on the distribution of the initial conditions at the $n$-th iteration stage; in particular, if the controls are computed on-the-fly, the controls will actually be feedback controls depending on the current state and the time elapsed since $t=0$.

@@

\recheck{
We derived linear response formulas for nonequilibrium molecular
dynamics (driven Langevin dynamics), which is a generalization of the
standard finite-time equilibrium response theory.  The novel response theory
dose not rely on any assumption on the existence of the invariant
measure nor stationary state. We validated the formula in numerical
experiments in comparison to brute-force nonequilibrium
simulations. There, we demonstrated that the nonequilibrium linear
response formula allows to extend the algorithmic use of linear
response theory to significantly stronger perturbations of the system
since it permits intermediate drivings, with which the effective perturbations
still lie in the linear range.
}

>>>>

\paragraph*{Remark.}
\recheck{Ciccotti and Jacucci have pointed out that
  the equilibrium linear response
can also be calculated by a direct nonequilibrium simulation of the
perturbed process, and  the nonequilibrium
approach has a smaller statistical uncertainty in comparison with the
standard Green-Kubo fomula~\cite{ciccotti1975direct}.
}

@@

\subsubsection*{Girsanov transformation}
We will briefly review the idea of the change of drift via Girsanov transformations; for details we refer to the textbook \cite{oksendal2003stochastic}. To this end, let $x_{t}=x_{t}(\omega)$ and $X_{t}=X_{t}(\omega)$ be the solutions of our generic stochastic differential equations
\begin{subequations}\label{sde2}
\begin{align}
dx_{t} & = b(x_{t},t)dt + a(x_{t})dB_{t}\,,\quad 0\le t\le T  \label{sde2-1}\\
dX_{t} & = (b(X_{t},t) + \eps v_{t})dt +a(X_{t})dB_{t}\,,\quad 0\le t\le T \label{sde2-2}
\end{align}
\end{subequations}
for $T<\infty$ and deterministic initial conditions
\[
x_{0}(\omega) = X_{0}(\omega) = x\quad\textrm{(almost surely)}\,.
\]
Suppose that there exists an auxiliary stochastic process $\xi_{t}\in\R^{n}$ such that 
\begin{equation}\label{xi}
a(X_{t})\xi_{t} = v_{t}\,.
\end{equation}
The auxiliary variable $\xi$ will be called \emph{control variable}. We define  
\[
W_{t} = \eps\int_{0}^{t}\xi_{s}\,ds + B_{t}\,,\quad 0\le t\le T\,,
\]
which allows us to rewrite (\ref{sde2-2}) as
\begin{equation}\label{sde2-3}
dX_{t} = b(X_{t},t)dt +a(X_{t})dW_{t} 
\end{equation}
It follows from the Girsanov theorem \cite[Thm.~8.6.8]{oksendal2003stochastic}, sometimes also called \emph{Cameron-Martin-Girsanov theorem} \cite{stroock2006}, that $W_{t}$ is again a Brownian motion under a new probability measure that has a density \wrt the Gaussian probability measure that is generated by the Brownian motion $B_{t}$. Specifically, let $P$ denote the law of the Brownian motion $B_{t}$ and define a new probability measure $Q$ on the space of continuous trajectories by 
\[
dQ=M_{T}dP
\]
with 
\begin{equation}\label{likelihood}
M_{t} = \exp\left(-\eps\int_{0}^{t}\xi_{s}\cdot dB_{s} - \frac{\eps^{2}}{2}\int_{0}^{t}|\xi_{s}|^{2}ds\right)\,,\quad 0\le t\le T\,.
\end{equation}
Technical details aside, the Girsanov theorem implies that $W_{t}$  for any functional 
\[
\varphi(\omega)=\int_{0}^{T}f((X_{s}(\omega))\,ds + g(X_{T}(\omega))
\] 
that is integrable \wrt $\nu$, we have the identity\footnote{One of the omitted technical details is Novikov's condition \cite[pp.~162]{oksendal2003stochastic} that guarantees that $(M_{t})_{0\le t\le T}$ is a Martingale, which by $\bE_{Q}[1]=\bE_{P}[M_{T}]=1$ implies normalizability of the new probability measure $Q$.}
\[
\bE_{P}[\varphi] := \int_{\Omega} \varphi(\omega)\, dP(\omega) = \int_{\Omega} \varphi(\omega) M_{T}^{-1}(\omega)\,dQ(\omega) =: \bE_{Q}[M^{-1}_{T}\varphi]\,.
\]
where 
\[
M_{T}^{-1} = \exp\left(\eps\int_{0}^{T}\xi_{s}\cdot dW_{s} - \frac{\eps^{2}}{2}\int_{0}^{T}|\xi_{s}|^{2}ds\right)
\]
is the density of $P$ relative to $Q$. Note that the expectation on the right hand side corresponds to the unperturbed dynamics, because $W_{t}$ is a standard Brownian motion under $Q$, and the expectation is over all realizations of (\ref{sde2-3}) starting from $X_{0}=x$. On the other hand, $X_{t}$ under $P$ corresponds to the perturbed dynamics, which should become clear upon comparing equations (\ref{sde2-2}) and (\ref{sde2-3}).

@@

We are interested in the dynamics under a small perturbation $\delta u$ of the control $u$, Specifically, we assume that $v$ in (\ref{sde2}) is of the form 
\begin{equation}\label{eqn:langevin-1c}
v_{t} = c(X_{t})\delta u(t)
\end{equation}
so that equation (\ref{xi}) that determines the change of measure in terms of the auxiliary control variable $\xi$ (and thus the linear response) reads 
\[
\sqrt{2\beta^{-1}}{\gamma}^{1/2} \xi_{t} = D(X_{t})\,\delta u(t)\,.
\]
Hence the equation for $\xi_{t}$ is solvable if and only if $\mathrm{range}(D(\cdot))\perp \mathrm{ker}(\gamma^{1/2})$, which, since $\gamma$ has full rank, means no restriction on $D(\cdot)$.
  % means that $D(\cdot)$ must be invertible almost everywhere.}

@@

The nonequilibrium linear response formula can be used for solving certain optimal control problems. To this end, let us remain in the setting of equations (\ref{eqn:langevin-1d})--(\ref{eqn:neq-response}) and assume that we are interested in choosing the nonequilibrium forcing $u$, such that the expected value $I=\bE[\psi]$ of some utility function 
\[
\psi(u) = \int_{0}^{T}\left\{f(x_{s}) - \frac{1}{2}|u(s)|^{2}\right\}\,dt + g(x_{T}) 
\]
is maximized where $x_{t}$ is the solution to the controlled Langevin equation (\ref{eqn:langevin-1d}) for $\eps=0$; the functions $f$ and $g$ are the running cost and the terminal cost, which are assumed to be continuous and bounded from above; without loss of generality, $f$ and $g$ are assumed to depend only on the positions $q:=x_{1}$ and not on the momenta $p:=x_{2}$. The quadratic term is a penalization that makes sure that the optimal control remains bounded \cite{schuette2012,hartmann2012}. Let us moreover assume that the controls are open-loop (i.e. without feedback) and can be represented by 
\[
u(t) = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
\]
with suitably chosen time dependent, bounded and Lipschitz continuous vector fields $\Phi_k\colon[0,T]\to\R^{n}$, $k=1,\ldots,K$. %In principle we could even go for closed-loop (i.e.~feedback) controls by making $\Phi_{k}$ state dependent. 
Specifically, we want to solve the optimal control problem of the following standard form: 
\begin{equation}\label{ocp1}
\begin{aligned}
 \max_{u\in\cU} \,& I(u)\quad \textrm{s.t.}\\
dq_t & =  \nabla_p H(q_t,p_t)dt\\
dp_t & =  -\nabla_q H(q_t,p_t)dt - \gamma \nabla_p H(q_t,p_t)dt + D(q_t)u(t) dt+\sigma dB_t\,,
\end{aligned}
\end{equation}
with $(q_0, p_0)$ being subject to a given initial distribution $\mu$, and
\begin{equation}\label{ocp2}
I(u) = \bE\left[\int_{0}^{T}\left\{f(q_{s}) - \frac{1}{2}|u(s)|^{2}\right\}\,dt + g(q_{T}) \right]\,,
\end{equation}
with the expectation being over all realizations of the controlled Langevin dynamics in (\ref{ocp1}) with the prescribed initial distribution. Here, the space of admissible controls $\cU$ consists of all bounded controls of the form 
\begin{equation}\label{admissible}
u(t) = \sum_{k=1}^K a_k \Phi_k(t)\,,\quad a_{k}\in \R\,,
\end{equation}

@@

In principle, optimal control problems such as (\ref{ocp1})--(\ref{ocp2}) can be solved by dynamic programming, i.e.~by solving the corresponding Hamilton-Jacobi-Bellman PDE \cite{fleming2006}. Except for very simple, essentially one-dimensional systems, solving Hamilton-Jacobi-Bellman equations is not an easy task, so we pursue a different strategy here. The idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and that we can do a gradient search in the unknown coefficients $a_{k}$, using the iteration
\[
u^{(n+1)} = u^{(n)} + \tau_{n}\nabla I(u^{(n)})\,,
\]
with $\tau_{n}>0$ being an adjustable parameter that determines the length of each gradient step. The gradient of $I$ can be easily evaluated using the linear response formula. To see this recall the notion of functional (G\^ateaux) derivatives as directional derivatives along a function $v$ (from a suitable function space): 
\begin{equation*}
\frac{\delta I}{\delta u} = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps v) = \langle\nabla I(u),v\rangle\,.
\end{equation*} 
For existence of the functional derivative \wrt general It\^o stochastic processes, we refer to \cite{bell2006}.  
Now the idea is to use that we have restricted our space of admissible controls to functions of the form (\ref{admissible}) with given basis vector fields and do a gradient search in the unknown coefficients $a_{k}$. This requires to compute the gradient \wrt the coefficients. Let the vector  
\[
\delta u = \sum_{k=1}^K \delta a_k \Phi_k(t)
\]   
denote the direction along which we want to differentiate where $\delta a_{1},\ldots,\delta a_{K}$ are the coefficients of the vector $\delta u$ in the basis of the $\Phi_{k}$, and note that 
\begin{equation*}
\begin{aligned}
\frac{\delta I}{\delta u} & = \left.\frac{d}{d\eps}\right|_{\eps=0} I(u+\eps \delta u)\\
& = \left(\lim_{\eps\to 0}\frac{I(u+\eps\delta u) - I(u)}{\eps}\right)\cdot \delta u\\
& = \left(\lim_{\eps\to 0}\frac{\bE_{P^{\eps}}[\psi(u+\eps \delta u)] - \bE_{P^{0}}[\psi(u)]}{\eps}\right)\cdot\delta u\\
\end{aligned}
\end{equation*} 
provided that the limit exists. The above iteration therefore is equivalent to
\begin{equation}\label{gd}
a_{k}^{(n+1)} = a_{k}^{(n)} + \tau_{n}\left.\frac{\partial I}{\partial a_{k}}\right|_{a_{k}=a_{k}^{(n)}}\,,
\end{equation}
with
\begin{equation}\label{dida}
\frac{\partial I}{\partial a_{k}} = \bE_{P^0}\left[\psi\int_{0}^{T}(\sigma^{-1}D(q_{s})\Phi_{k}(s))\cdot dB_{s} + \int_{0}^{T}u(s)\cdot\Phi_{k}(s)\,ds \right] \,,
\end{equation}
and $\bE_{P^0}[\cdot]$ being the expectation over all realizations of (\ref{eqn:langevin-1d}) with $\eps=0$ and $u=u^{(n)}$ being the current iterate and arbitrary initial conditions $x=(q,p)$, e.g., 
\[
\bE_{P^{0}}[\psi] = \int_{\R^{2n}}\bE_{P^{0}_{x}}[\psi]\,d\mu(x) \,,
\]
with $P_{x}^{0}$ denoting the corresponding path measure with deterministic initial condition, and $\mu$ is the prescribed initial distribution.
\recheck{A related result on computing the parameter sensitivity can be found
  in Ref.~\cite{warren2012malliavin}.}
Note that we can compute all partial derivatives $\partial I/\partial a_k$, $k=1,\ldots,K$ from just one ensemble of nonequilibrium paths of $(q_{t},p_{t})$. Further note that, in general, the updated coefficient $a^{(n+1)}$ and hence the updated control $u^{(n+1)}$ will depend on the distribution of the initial conditions at the $n$-th iteration stage; in particular, if the controls are computed on-the-fly, the controls will actually be feedback controls depending on the current state and the time elapsed since $t=0$.

@@

\recheck{
  We derived the first and second order
  response formulas for nonequilibrium molecular
dynamics (driven Langevin dynamics), which is a generalization of the
standard finite-time equilibrium linear response theory.
The novel nonequilibrium response theory
does not rely on any reversibility or stationarity assumption on the dynamics.
We validated the formula in numerical
experiments in comparison to brute-force nonequilibrium
simulations. There, we demonstrated that the nonequilibrium linear
response formula allows to extend the algorithmic use of linear
response theory to significantly stronger perturbations of the system
since it permits intermediate drivings, with which the effective perturbations
still lie in the linear range.
}

### 2013-08-23 writedown/linear.response/reply/reply.tex
improves according to Carsten

We have rewritten Sec. 2.1, and restricted the discussion of the
equilibrium response theory (3) only in the finite-time
horizon, so that it is comparable to the the nonequilibrium response theory (10), which is also
developed in the finite-time horizon. Our work
can be treated as a
generalization of the equilibrium response theory in the sense that the
reference process is not assumed to be in equilibrium nor in stationary.
The introduction and conclusion are revised accordingly.

@@

As far as we concerned, we do not have a clear answer to this question
in the most general case.
However, in this paper, we deal with  the Langevin dynamics.
The friction
matrix is positive definite, so the auxiliary process $\xi_t$ is uniquely
solvable.
\\

@@

In the paper by Giovanni and Jacucci, the authors use direct
nonequilibrium molecular dynamics simulation to 
calculate the \emph{equilibrium response}, i.e. the mobility of the charged
Lennard-Jones particle, and show that
the nonequilibrium approach has the advantage of smaller statistical error over
the Green-Kubo formula. Our  nonequilibrium simulations,
in comparison, calculate the
\emph{nonequilibrium response}.
As far as the authors concern, the method presented in the paper is new and
has never been studied.

>>>>

We have rewritten Sec. 2.1, and restricted the discussion of the
equilibrium response theory (3) only in the finite-time
horizon, so that it is comparable to the nonequilibrium response theory (10), which is also
developed in the finite-time horizon. Our work
can be treated as a
generalization of the equilibrium response theory in the sense that the
reference process is not assumed to be in equilibrium.
Being more general, our theory does not assume the existence of nonequilibrium steady state.
The introduction and conclusion are revised accordingly.

@@

% As far as we concerned, we do not have a clear answer to this question
% in the most general case.
In this paper, when we deal with  the Langevin dynamics, 
the friction
matrix is positive definite, so the auxiliary process $\xi_t$ is uniquely
solvable.
If the matrix in front of the noise does not have full rank, then
there are some components of the driving Brownian motion in the SDE
that are uncorrelated from $X_t$, in which case the extra terms are
independent of $\varphi$ that depends only on $(X_t)$. Hence the
corresponding 1st and 2nd order linear response expressions vanish
because of the Martingale property $\textbf{E}[\int \xi^{extra}\cdot dB^{extra}]=0$
and the Ito isometry $\textbf{E}[(\int \xi^{extra}\cdot dB^{extra})^2]= \textbf{E} [\int
|\xi^{extra}|^2dt]$.
\\

@@

In the paper by Giovanni and Jacucci, the authors use direct
nonequilibrium molecular dynamics simulation to 
calculate the \emph{equilibrium response}, i.e. the mobility of the charged
Lennard-Jones particle, and show that
the nonequilibrium approach has the advantage of smaller statistical error over
the Green-Kubo formula. Our  nonequilibrium simulations,
in comparison, calculate the
\emph{nonequilibrium response}.
As far as the authors concern, the method presented in the paper is new and
has never been studied. The citation and discussion to Giovanni and Jacucci's
work have been added to the manuscript.
